---
title: "Monitoria 1 - Métodos Computacionais em Finanças"
author: "Matheus Carrijo de Brito"
date: "2022-20-07"
output: 
  prettydoc::html_pretty:
    bibliography: biblio.bib
    theme: cayman
    hightlight: github
    math: katex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
# Importing some packages 

rm(list = ls())

library(quantmod)
library(quantreg)
library(tidyquant)
library(ggplot2)
library(data.table)
#library(fOptions)
library(RQuantLib)
library(fAssets)
library(tidyverse)

```

# Introdução

## Processo estocastico de wiener

It is a specific case of a markovian stochastic process, also called a Brownian Movement. Let z be a RV that follows this process, then it is characterized by the followings properties:

1 - The change of the process $\Delta z$ in a small amount of time $\Delta t$ is given by:

$$
\Delta z = \epsilon \sqrt{\Delta t} \hspace{2cm} \epsilon ~ N(0, 1)
$$

2 - The changes in z are independent for any interval of time

$$
\Delta z_{t} \perp \!\!\! \perp \Delta z_{j} \hspace{2cm} \forall t,j, \quad t\neq j 
$$ For a long space of time the process Z is the sum of small pieces that are distributated as independent normal variables, which leads to z be normaly distributed

$$
z(T) - Z(0) = \sum^{N}_{t = 1} \epsilon_{t}\sqrt{\Delta t} \sim N(0, T)
$$

## Generalized Wiener process

In general we are interest a linear function of a wiener process given by:

$$
\Delta x = \alpha  \Delta t + \beta  \Delta z  \hspace{2cm} \{\alpha, \beta\} \in \mathbb{R}
$$ The parameter $\alpha$ gives the drift variation in the process (taxa de deriva), the parameter $\beta$ is the noise in the process (taxa de variância). It is easy to note that the process is normaly distributed

$$
X(T) - X(0) \sim N(\alpha T, \beta^{2} T)
$$

## Ito process

This process generalize the wiener process by allowing that the coefficients $\alpha, \beta$ be function of the underline variables $x, t$. Thus, the process is given by:

$$
\Delta x = \alpha(x, t) \Delta t + \beta(x,t) \Delta z
$$ Example: The path of a stock is often modeled as a simple geometric browninan movement $\frac{\Delta S}{S} = \mu \Delta t + \sigma \epsilon \sqrt{\Delta t}$, thus we already now the distribution

$$
\frac{\Delta S}{S} = \mu \Delta t + \sigma \epsilon \sqrt{\Delta t} \sim N(\mu \Delta t, \sigma^{2} \Delta t)
$$

## Let's simulate and analyse the model below

First, let generate random numbers

```{r}
# Repeat the same aleatory numbers 

set.seed(123)

years <- 4

epsilons <- rnorm(252*years, mean = 0, sd = 1/252)

# Setting some parameters and initial conditions 

P_0 <- 100 
mu = 0.10
sigma = 0.25
delta_t <- 1/252


```

# Process 1:

-   $\Delta P_{t+1} = \mu \Delta t + \sigma N(0, \Delta t)$

Is given by a browning motion with constant drift and noise. The simulation is reported below.

```{r, fig.align="center"}

P <- c(P_0, rep(0, length = length(epsilons) -1))

for(i in 1:(length(epsilons)-1)){
  
  P[i+1] <- P[i] +  mu * delta_t + epsilons[i] * sigma
  
}

# Plotting the Asset Price 

ggplot() + 
  geom_line(aes(x = 1:length(epsilons), y = P)) + 
  scale_x_continuous(breaks = seq(1,1001,100)-1) +
  labs(x = "Time", y = "Asset Price")


# let's make a distribution 

P_0 = 100
mu = 0.1
sigma = 0.25
delta_t = 1/252
n_simulation = 1000
lenght_asset = 1000



```

### Making the simulation function

```{r, warning=F}
simulations <- function(P_0 = 100, mu = 0.1, 
                        sigma = 0.25, delta_t = 1/252,
                        n_simulation = 1000, lenght_asset = 1000){
  
  # Creating a matrix to save the simulations  
  sim_vector <- matrix(0, nrow = lenght_asset, ncol = n_simulation)
  
  for(j in seq_len(n_simulation)){
    
    # Create an empty vector to simulate the asset price path 
    P <- c(P_0, rep(0, length = length(lenght_asset)-1))
    
    # Draw the shock 
    epsilons <- rnorm(n_simulation, mean = 0, sd = delta_t)
    
    # Make a simulation  
    for(i in 1:(n_simulation)){
    
    P[i+1] <- P[i] +  mu * delta_t + rnorm(n = 1, mean = 0, sd = delta_t) * sigma
    
    # Save the results 
    sim_vector[i, j] <- P[i]
    
    }
    
  }  
  
  # Returning the matrix
  return(data.frame(sim_vector))
}
```

## Making the simulation and plotting

```{r, warning=F}
sim <- simulations(P_0 = 100, mu = 0.1, 
                   sigma = 0.25, delta_t = 1/252, 
                   n_simulation = 1000, lenght_asset = 1000)


# Plotting the first 50 simulations 
sim[, 1:50] |> 
  mutate(n = seq_len(1000)) |>
  melt(id.vars = "n", variables.name = "variable", value.name = "value") |>
  group_by(variable) |> 
  ggplot() + 
  geom_line(aes(x = n, y = value, colour = variable)) + 
  scale_color_grey() +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  labs(x = "Time", y = "Asser Price", title = "Simulations")


# Making anothers simulation changing parameters 

simulations(P_0 = 100, mu = 1, 
                   sigma = 2,
                   delta_t = 1/252, 
                   n_simulation = 1000, 
            lenght_asset = 1000) |> 
  select(1:50) |> 
  mutate(n = seq_len(1000)) |>
  melt(id.vars = "n", variables.name = "variable", value.name = "value") |>
  group_by(variable) |> 
  ggplot() + 
  geom_line(aes(x = n, y = value, colour = variable)) + 
  scale_color_grey() +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5, face = "bold")) + 
  labs(x = "Time", y = "Asser Price", title = "Simulations")



```

# Process 2

-   $\Delta P_{t+1} = \mu P_{t} \Delta t + \sigma P_{t} N(0, \Delta t)$

This process is a geometric browning motion that is often used to model the stock price movements, since the variation is relative to the initial point of the stock.

```{r , fig.align="center"}

# Function that iterates the process 2

process_2 <- function(p, n_path, mu, delta_t, sigma, p_0){

  
  p[[1]] <- p_0
  
  for(i in seq_len(n_path)){
  
  p[[i+1]] <- p[[i]]*(1 + mu*delta_t + sigma*10*rnorm(n =1, mean = 0, sd = delta_t))
  
  }
  
  return(p) 
  
}

# Create an empty list where the simulation will be stored
n_path = 100
n_sim = 100

p <- vector(mode = "list", n_sim)

# Another way to run n simulations  
 
process_2_df <- do.call("rbind",  
  lapply(p, function(x) process_2(p = x, n_path = n_path, mu = 0.1, delta_t = 1/252, sigma = 0.20, p_0 = 100))) |> 
  as.data.frame() |> 
  setNames(c("sim", seq_len(n_sim))) 


# A simple way to plot the paths  
matplot(process_2_df |> t(), type = "l", ylab= "Asset Price", xlab = "Time")

# Plotting all simulations of the Asset Price path 
process_2_df |> 
  t() |>
  data.frame() |>
  mutate(n = 1:(n_sim+1)) |>
  gather(-c("n"), key = "variable", value = "value") |>
  mutate(value = as.numeric(value)) |> 
  group_by(variable) |> 
  ggplot() +
  geom_line(aes(x = n, y = value, colour = variable)) +
  theme(legend.position = "none") +
  labs(x = "Time", y = "Asset Price") + 
  scale_colour_grey()

```

### Process 3:

-   $\Delta P_{t+1} = \mu(\bar{P} - P_{t}) \Delta t + \sigma P_{t} N(0, \Delta t)$

This process is a geometric browning motion with normalized drift. Note that for a price level higher than $\bar{P}$ the drift becomes negative and the process start to converge to the mean. The inverse happens when $P_{t} > \bar{P}$. This process may not be acceptable to model stock prices, since its expected an positive growth. But it may be adequate to model interest rate since its a stationary variable.  

```{r, fig.align="center"}

process_3 <- function(n_sim = 100, time_length = 100, p_0 = 100, 
                      mu = 0.1, delta_t = 1/252, sigma= 0.2){
  
  dt_process_3 <- matrix(NA, nrow = time_length, ncol = n_sim) 
  
  dt_process_3[1, ] <- p_0
  
  for(j in seq_len(n_sim)){
  
    for(i in seq_len(time_length-1)){
      
      dt_process_3[i+1, j] <- dt_process_3[i, j] + 
        mu*(mean(dt_process_3[, j], na.rm = T) - dt_process_3[i, j])*delta_t + 
        dt_process_3[i, j]*sigma*rnorm(n = 1, mean = 0, sd = delta_t)
      
    }
    
  }
  
  dt_process_3 <- data.frame(dt_process_3)
  return(dt_process_3)
}

dt_3 <- process_3(n_sim = 100)


# Plotting Asset Price Path 

dt_3 |> 
  mutate(sim = seq_len(100)) |> 
  tidyr::gather(-c("sim"), key ="variable", value = "value") |>
  group_by(variable) |> 
  mutate(high = ifelse(sim == 50 & value > 100,1,0), 
         high = max(high)) |>
  group_by(variable, high) |>
  ggplot() + 
  geom_line(aes(x = sim, y = value, group = variable, colour = as.factor(high))) + 
  labs(x = "Time", y = "Asset Pricing") + 
  scale_colour_manual(values = c("grey", "black"), 
                     labels = c("Below 100 in time 50", "Above 100 in time 50"),
                     name = "Legend") + 
  theme(legend.position = "bottom")


```

### Process 4:

-   $\Delta P_{t+1} = \mu P_{t} \Delta t + \sigma P_{t} N(\Delta P_{t}/P_{t-1}, \Delta t)$

This process have the same drift of a Geometric Brownian motion, the disturbance that is different. When returns are positive the shock becomes, on average, positive, implying in an increase in the asset price. 

```{r, fig.align="center"}

length_path <- 100  # Number of steps in asset price simulation 
n_sim <- 100 # Number of Simulation 

# Parameters of the simulations 

mu <- 0.1
delta_t <- 1/252
sigma <- 0.4


# Making the simulation 

process_4 <- matrix(data = NA, nrow = length_path, ncol = n_sim)

# Filling the first and second rows with initial Price P_0 

process_4[1:2, ] <- P_0 

for(j in seq_len(n_sim)){
  
  for(i in seq(from = 3, to = n_sim, by = 1)){
    
    process_4[i, j] <- process_4[i-1, j]*(1 + mu * delta_t + sigma*rnorm(n = 1, mean = (process_4[i-1]- process_4[i-2])/process_4[i-2], sd = delta_t))
    
  }
}

matplot(y = process_4, type = "l", xlab =  "Time", ylab = "Aset Price")

# Plotting Asset Price path 

process_4 |> 
  data.frame() |> 
  mutate(time = seq(from = 1, to = length_path, by = 1)) |>
  gather(-c("time") , key = "simulation", value = "value") |>
  ggplot() + 
  geom_line(aes(x = time, y = value, colour = simulation)) +
  theme(legend.position = "none") +
  labs(x = "Time", y = "Asset Price") + 
  scale_colour_grey()

```

### Process 5:

-   $\Delta P_{t+1} = \mu( \bar{P} - P_{t}) \Delta t + \sigma P_{t} N_{1}(0, \Delta t) + \sigma_{j}Bern(p)P_{t} N_{2}(0, \Delta t)$ where, $N_{1}, N_{2}$ are independent and $Bern(p)$ is a Bernoulli with probability $p$ at the interval $\Delta_{t}$

This process is crazy with jumps

```{r, fig.align="center"}


length_path <- 100  # Number of steps in asset price simulation 
n_sim <- 100 # Number of Simulation 

# Parameters of the simulations 

mu <- 0.1
delta_t <- 1/252
sigma <- 0.4
P_0 <- 100


# Making the simulation 

process_5 <- matrix(data = NA, nrow = length_path, ncol = n_sim)

process_5[1, ] <- P_0

# Making the simulation 

for(j in seq_len(n_sim)){

  for(i in seq(from = 2, to = length_path, by = 1)){
    
    process_5[i, j] <- process_5[i-1, j] + mu*(mean(process_5, na.rm = T) - process_5[i-1, j]) * delta_t + 
      process_5[i-1, j]*sigma*rnorm(n = 1, mean = 0, sd = delta_t) + 
      sigma*rbinom(n = 1 , size = 1, prob = 0.5)*rnorm(n = 1, mean = 0, sd = delta_t)
    
  }
  
}


# Plotting Asset Price path 

process_5 |> 
  data.frame() |> 
  mutate(time = seq(from = 1, to = length_path, by = 1)) |>
  gather(-c("time") , key = "simulation", value = "value") |>
  ggplot() + 
  geom_line(aes(x = time, y = value, colour = simulation)) +
  theme(legend.position = "none") +
  labs(x = "Time", y = "Asset Price") + 
  scale_colour_grey()

```
