---
title: "computation_methods_ta_2"
author: "Matheus Carrijo de Brito"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    theme: paper
    highlight: textmate
    number_sections: true
    df_print: paged
    code_folding: hide # Option to add hide button in Rmarkdown
    bibliography: biblio.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Packages

```{r message=FALSE, warning=FALSE, include=FALSE}
# Importing some packages 

rm(list = ls())

library(quantmod)
library(quantreg)
library(tidyquant)
library(ggplot2)
library(data.table)
library(RQuantLib)
library(fAssets)
library(tidyverse)

```


# Expectation using Monte Carlo 

## Let's assume that stock prices follows a GBM 

$$
\Delta P_{t+1} = \mu P_{t} \Delta t  + \sigma P_{t}N(0, \Delta t) \hspace{1cm} \text{with}\hspace{1cm} \mu = 0.1, \sigma = 0.5 
$$
## Making the simulations 

#### Writing the simulation function 

```{r}
simulations <- function(P_0 = 100, mu = 0.1, 
                        sigma = 0.5, delta_t = 1/252,
                        n_simulation = 1000, lenght_asset_path = 252){
  
  # Creating a matrix to save the simulations  
  matrix_simulation <- matrix(0, nrow = lenght_asset_path, ncol = n_simulation)
  
  # Setting the columns names 
  
  colnames(matrix_simulation) <- paste("sim", seq_len(n_simulation), sep = "_")
  
  # Filling the first row with inittial price
  matrix_simulation[1, ] <- P_0
  
  # In each simulation we generate the asset price path 
  # seg_len generates a sequence from 1, to n_simulation, by = 1
  
  for(j in seq_len(n_simulation)){

    # Make a simulation  
    # 2:(lenght_asset_path) is a sequence that start at the second row (time) and ends at the last row (time) 
    for(i in 2:(lenght_asset_path)){
    
    matrix_simulation[i, j] <- matrix_simulation[i-1, j] *(1 +  mu * delta_t +  sigma * 10 * rnorm(n = 1, mean = 0, sd = delta_t))
    
    } # End of the asset price loop 
    
  } # End of the simulation loop 
  
  # Returning the matrix as data frame 
  return(data.frame(matrix_simulation))
}
```

#### making the simulations

```{r cols.print = 100, rows.print = 10}

sim_prices <- simulations(P_0 = 50, mu = 0.1, sigma = 0.25, delta_t = 1/252, n_simulation = 100,
                          lenght_asset_path = 252)

# Printing the first 10 rows and 5 columns 
kableExtra::kable(x = sim_prices[1:10, 1:5], format = "html") |>
  kableExtra::kable_styling()


```

Given that the price of the stock is R$ 50, answer the following questions:

- What is the probability that the price of the stock will be above R$ 70 after one year?

```{r}

# sim_prices[252, ] Gives the values of the stock price after one year for all simulations 

# sim_prices[252, ] > 70 compute the logical condition for each simulation the stock price is above 70 and return a logical condition 

# Sum this logical vector returns all True condition 

# Dividing by the number of columns gives the probability 
 
sprintf("Probability: %s", sum(sim_prices[252, ] > 70)/ncol(sim_prices))

```

- What is the probability that the price of the stock will be above R$ 70 at some point in the next year?

```{r}

prob <- c()

for(j in seq_along(sim_prices)){
  
  prob[j] <- sum(sim_prices[, j] > 70) >= 1
  
}


sprintf("Probability:  %s", sum(prob)/ncol(sim_prices))

# Visualization of the paths 

sim_prices |> 
  mutate(time = seq_len(nrow(sim_prices))) |>
  gather(-c("time"), key = "simulation", value = "value") |> 
  group_by(simulation) |> 
  mutate(high = ifelse(value > 70, 1,0), 
         high = max(high)) |> 
  ungroup() |> 
  ggplot() + 
  geom_line(aes(x = time, y = value, colour = as.factor(high), group = simulation)) +
  scale_color_manual(values = c("red", "blue"), name = "Legend", labels = c("UP", "Down")) + 
  theme(legend.position = "bottom")

```

- What is the probability that the price of the stock will be above R\$ 70 after one year and that at some point during the year the price reached below R\$ 50?

```{r}

prob <- c()

for(j in seq_along(sim_prices)){
  
  prob[j] <- sum(sim_prices[, j] < 50 & sim_prices[nrow(sim_prices), j] > 70) >= 1
  
}

sprintf("Probability:  %s", sum(prob)/ncol(sim_prices))

# Visualization of the paths that satisfies the conditions above 

sim_prices |> 
  mutate(time = seq_len(nrow(sim_prices))) |>
  gather(-c("time"), key = "simulation", value = "value") |> 
  group_by(simulation) |> 
  mutate(high = ifelse(value < 50 & value[time == 252] > 70, 1,0), 
         high = max(high)) |> 
  ungroup() |> 
  ggplot() + 
  geom_line(aes(x = time, y = value, colour = factor(high), alpha = factor(high), group = simulation)) +
  scale_color_manual(values = c("grey", "black"), name = "Legend", labels = c("Down", "UP")) +
  scale_alpha_manual(values = c(0.5,1)) + 
  theme(legend.position = "bottom")

```

Another advanced way 

```{r}
sprintf("Probability:  %s",  
  apply(sim_prices, FUN = function(x) max(x >= 70), MARGIN =  2) |> sum()/ncol(sim_prices)
)

```

- What is the probability that the price of the stock will be above R\$ 70 after one year if during the first six months the price was below R\$ 70?

```{r}

prob <- c()

for(j in seq_along(sim_prices)){
  
  prob[j] <- (sum(sim_prices[1:151, j] < 70) == 151 & # If all price are bellow 70 the sum needs to be equal 151
                sum(sim_prices[nrow(sim_prices), j > 70])) >= 1 
  
}

sprintf("Probability:  %s", sum(prob)/ncol(sim_prices))


# Visualization of the paths that satisfies the conditions above 

sim_prices |> 
  mutate(time = seq_len(nrow(sim_prices))) |>
  gather(-c("time"), key = "simulation", value = "value") |> 
  mutate(first_six_months = ifelse(time <= 151,1,0)) |> 
  group_by(simulation) |> 
  mutate(high = ifelse(first_six_months == 1 & value < 70 & value[time == 252] > 70, 1,0)) |> 
  ggplot() + 
  geom_line(aes(x = time, y = value, colour = as.factor(high), group = simulation)) +
  scale_color_manual(values = c("red", "blue"), name = "Legend", labels = c("Down", "UP")) + 
  theme(legend.position = "bottom")


```

- What is the probability that the price of the stock crosses three times the R\$ 70 threshold from below in one year?


```{r}

prob <- c()

for(j in seq_along(sim_prices)){
  
  prob[j] <- sum(sim_prices[, j] > 70) >= 3
  
}

sprintf("Probability:  %s", sum(prob)/ncol(sim_prices))


# Visualization of the paths that satisfies the conditions above 

sim_prices |> 
  mutate(time = seq_len(nrow(sim_prices))) |>
  gather(-c("time"), key = "simulation", value = "value") |> 
  group_by(simulation) |> 
  mutate(high = ifelse(sum(value > 70) >= 3, 1,0), 
         high = max(high)) |> 
  ggplot() + 
  geom_line(aes(x = time, y = value, colour = as.factor(high), group = simulation)) +
  scale_colour_manual(values = c("red", "blue"), name = "Legend", labels = c("Down", "UP")) + 
  theme(legend.position = "bottom")


```


