---
title: "Monitoria 1 - Métodos Computacionais em Finanças"
author: "Matheus Carrijo de Brito"
date: "2022-20-07"
output: 
  prettydoc::html_pretty:
    bibliography: biblio.bib
    theme: cayman
    hightlight: github
    math: katex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
# Importing some packages 

library(quantmod)
library(quantreg)
library(tidyquant)
library(ggplot2)
library(data.table)
#library(fOptions)
library(RQuantLib)
library(fAssets)

```

## Introdução 

### Processo estocastico de wiener 

It is a specific case of a markovian stochastic process, also called a Brownian Movement. Let z be a RV that follows this process, then it is characterized by the followings properties: 

1 - The change of the process $\Delta z$  in a small amount of time $\Delta t$ is given by: 

$$
\Delta z = \epsilon \sqrt{\Delta t} \hspace{2cm} \epsilon ~ N(0, 1)
$$

2 - The changes in z are independent for any interval of time 

$$
\Delta z_{t} \perp \!\!\! \perp \Delta z_{j} \hspace{2cm} \forall t,j, \quad t\neq j 
$$
For a long space of time the process Z is the sum of small pieces that are distributated as independent normal variables, which leads to z be normaly distributed 

$$
z(T) - Z(0) = \sum^{N}_{t = 1} \epsilon_{t}\sqrt{\Delta t} \sim N(0, T)
$$
#### Generalized Wiener process  

In general we are interest a linear function of a wiener process given by: 

$$
\Delta x = \alpha  \Delta t + \beta  \Delta z  \hspace{2cm} \{\alpha, \beta\} \in \mathbb{R}
$$
The parameter $\alpha$ gives the drift variation in the process (taxa de deriva), the parameter $\beta$ is the noise in the process (taxa de variância). It is easy to note that the process is normaly distributed

$$
X(T) - X(0) \sim N(\alpha T, \beta^{2} T)
$$

#### Ito process  

This process generalize the wiener process by allowing that the coefficients $\alpha, \beta$ be function of the underline variables $x, t$. Thus, the process is given by: 

$$
\Delta x = \alpha(x, t) \Delta t + \beta(x,t) \Delta z
$$
Example: The path of a stock is often modeled as a simple geometric browninan movement $\frac{\Delta S}{S} = \mu \Delta t + \sigma \epsilon \sqrt{\Delta t}$, thus we already now the distribution 

$$
\frac{\Delta S}{S} = \mu \Delta t + \sigma \epsilon \sqrt{\Delta t} \sim N(\mu \Delta t, \sigma^{2} \Delta t)
$$

# Let's simulate and analyse the model below

First, let generate random numbers 

```{r}
set.seed(123)

epsilons <- rnorm(252*4, mean = 0, sd = 1/252)

# Setting some parameters and initial conditions 

P_0 <- 100 
mu = 0.10
sigma = 0.25

delta_t <- 1/252

```


### Process 1: 

- $\Delta P_{t+1} = \mu \Delta t + \sigma N(0, \Delta t)$

Is given by a browning motion with constant drift and noise. The simulation is reported below. 

```{r}

P <- c(P_0, rep(0, length = length(epsilons) -1))

for(i in 1:(length(epsilons)-1)){P[i+1] <- P[i] +  mu * delta_t + epsilons[i] * sigma}

ggplot() + 
  geom_line(aes(x = 1:length(epsilons), y = P)) + 
  scale_x_continuous(breaks = seq(1,1001,100)-1) +
  labs(x = "Time", y = "Asset Price")

```

## Process 2 

- $\Delta P_{t+1} = \mu P_{t} \Delta + \sigma P_{t} N(0, \Delta t)$

This process is a geometric browning motion that is often used to model the stock price movements, since the variation is relative to the initial point of the stock. 

```{r}

P_2 <- c(P_0, rep(0, length = length(epsilons)-1))

for(i in 1:(length(epsilons) - 1)){P_2[i+1] <- P_2[i]*(1 + mu*delta_t + sigma*10*epsilons[i])}

ggplot() +
  geom_line(aes(x = seq_along(epsilons), y = P_2)) +
  scale_x_continuous(breaks = seq(1,1000,100)-1) +
  labs(x = "Time", y = "Asset Price")

```

## Process 3: 

- $\Delta P_{t+1} = \mu(\bar{P} - P_{t}) \Delta t + \sigma P_{t} N(0, \Delta t)$

This process is a geometric browning motion with normalized drift.

```{r}

P_3 <- c(P_0, rep(NA, length = length(epsilons)-1))

for(i in 1:(length(epsilons) - 1)){P_3[i+1] <- P_3[i]+ mu*(mean(P_3, na.rm = T) - P_3[i])*delta_t + P_3[i]*sigma* epsilons[i+1]}

ggplot() + 
  geom_line(aes(x = seq_along(epsilons), y = P_3)) + 
  scale_x_continuous(breaks = seq(1,1001,100)-1) +
  labs(x = "Time", y = "Asset Pricing")

```

## Process 4: 

- $\Delta P_{t+1} = \mu  P_{t} \Delta t + \sigma P_{t} N(\Delta P_{t}/P_{t-1}, \Delta t)$

This process is crazy.  

```{r}

P_4 <- c(P_0, P_0, rep(NA, length(epsilons)-2))

for(i in 1:(length(epsilons)-2)){
  
  P_4[i+2] <- P_4[i+1]*(1 + mu * delta_t + sigma*rnorm(n = 1, mean = (P_4[i+1]- P_4[i])/P_4[i], sd = delta_t))
  
}

ggplot() + 
  geom_line(aes(x = seq_along(epsilons), y =  P_4)) +
  scale_x_continuous(breaks = seq(1,1001,100)-1) +
  labs(x = "Time", y = "Asset Price")

```

## Process 5: 

- $\Delta P_{t+1} = \mu( \bar{P} - P_{t}) \Delta t + \sigma P_{t} N_{1}(0, \Delta t) + \sigma_{j}Bern(p)P_{t} N_{2}(0, \Delta t)$ where, $N_{1}, N_{2}$ are independent and $Bern(p)$ is a bernoulli with probability $p$ at the interval $\Delta_{t}$

This process is crazy with jumps 

```{r}

P_5 <- c(P_0, rep(NA, length(epsilons) - 1)) 

for(i in seq_along(P_5)){
  
  P_5[i + 1] <- P_5[i] + mu*(mean(P_5, na.rm = T) - P_5[i]) * delta_t + 
    P_5[i]*sigma*rnorm(n = 1, mean = 0, sd = delta_t) + 
    sigma*rbinom(n = 1 , size = 1, prob = 0.5)*rnorm(n = 1, mean = 0, sd = delta_t)
  
}

ggplot() + 
  geom_line(aes(x = seq_along(P_5), y = P_5)) + 
  scale_x_continuous(breaks = seq(1,1001,100)-1) + 
  labs(x = "Time", y = "Asset Price")

```





